### 2013-11-15 CGP Journal/response.txt
Completed a first pass of addressing reviewer comments

In [8], the authors use AND, OR, NAND, NOR for all of the other boolean problems tested (Even Parity, Digital Adder, and Digital Comparitor),
but use AND, AND with one input inverted, XOR and OR on digital multiplier.
The two papers [8] cites that performed research on this problem used AND, XOR.  They claim this makes the problem easier as
human designed solutions use only AND, XOR.  Therefore eliminating XOR as a choice likely makes this problem even harder.

It is possible to achieve O(1) amortized cost when performing mutations using the probability method.
This can be done leveraging the binomial distribution, which in many languages is implemented to give
the random number of trials k of N total which are true given probability P.  k genes can then
be selected from the genome at random just as when using a mutation rate.  This operator is
advantageous in that allows for a flexible number of mutating genes, up to and including
mutating all genes in a single step.  If I were to extend Single to use a configurable
number of genes, this would be my choice of method.

Over four thousand runs would be required to properly tune and evaluate any additional
problem.
A major focus of our work (ordering effect on the location and number of active genes)
is grounded in theory which makes no assumptions on the type of problem being solved.
As the experimental results matched the theory, we have no reason to suspect our
conclusions on these topics to change when the operators or data type is changed.
Some conclusions that could change depending on data type: The relative effectiveness
of the proposed modifications (IE, skip vs single, reorder vs normal, etc), the
evolved genome sizes, and the charts in Table IV and Table V.

The reason we chose all of our original problems as binary was to allow for the
detailed analysis of behavior in the latter half of Section VI.  If we had done
four problems, but using diverse domains, it would have been much more challenging
to determine why we observed changes in behavior between problems.  By using
four problems with similar domains and the same operators, we are able to determine
how our techniques scale with problem difficulty.

While we agree it is desirable to have more problems, and more diverse problems,
we were limited by how deep we wanted to go in analyzing our set of problems.
As stated elsewhere, each problem required approximately four thousand runs to
determine optimal genome size and mutation rate.  Similarly we wanted to delve
deeply into behavior, which requires time to determine measures and discover patterns,
not all of which are going to be domain independent.  As CGP's foundation is the
binary domain, we felt this was the correct choice for such detail oriented work.

Upon the suggestion of Reviewer 2, we further examined the statistical measures
of evaluations to success which we report in Table II.  As it is not safe to assume the number
of evaluations to success is normally distributed (in fact, they are most likely not normal),
we chose to provide non-parametric (median) instead of parametric (mean) measures,
as was done in [8].  Upon our review, we decided Median Absolute Deviation was not
capturing useful information, and we now show bootstrap confidence intervals, which is a much
more statistically sound and interesting value.

As to the including the wall clock time for our experiments, we attempted to cover our
reasoning for not doing this in the 4rd to last paragraph in the Results section.
In order to provide a fair comparison of runtime, we would need to redo tuning
with a focus on minimizing wall clock time instead of number of evaluations, which
would require another four thousand runs to time just one problem.  An additional
problem in reporting this value is hardware diversity.  To complete our experiments
we utilized a computing cluster with diverse hardware quality.  To avoid this problem,
we would need to recreate all experiments on homogeneous hardware, which is intractable
for experiments of this size.