### 2014-10-15 Comprehensive/chapter-graybox.tex

\chapter{New Problem Domain: Gray Box Optimization}

\section{Problems in this domain}

\section{Formal Requirements}

\section{Efficient Local Search}

\section{Efficient Hamming Search}

\section{Memetic Algorithms}