### 2014-10-15 Comprehensive/chapter-proposed.tex

\chapter{Proposed Research}

\section{Simple Memetic Algorithm Using Efficient Search}

\section{Developing Efficient Linkage Aware Crossover}

\section{Comparison of Gray Box Methods}