### 2014-11-19 CV/cv.tex
Added Optimization to CV research interests, changed order

%______________________________________________________________________________________________________________________
% @brief    LaTeX2e Resume based on the example from Kamil K Wojcicki
\documentclass[margin,line]{resume}
\usepackage{verbatim}

%______________________________________________________________________________________________________________________
\begin{document}
\name{\Large Brian W. Goldman}
\begin{resume}

    %__________________________________________________________________________________________________________________
    % Contact Information
    \section{\mysidestyle Contact\\Information}

    6195 Gossard                            \hfill phone: (314) 313-1281          \vspace{0mm}\\\vspace{0mm}%
    East Lansing, Michigan, 48823                          \hfill email: brianwgoldman@acm.org          %\vspace{0mm}%\\\vspace{0mm}%
    %__________________________________________________________________________________________________________________
    % Research Interests
    \section{\mysidestyle Research\\Interests}
    Optimization, Artificial Intelligence, Evolutionary Computation, Machine Learning

    %__________________________________________________________________________________________________________________
    % Education
    \section{\mysidestyle Education}
    \textbf{Michigan State University}, East Lansing, Michigan \vspace{2mm}\\\vspace{1mm}%
    \textsl{Doctor of Philosophy in Computer Science \& Engineering} \hfill \textbf{ August 2012 -- Present}\vspace{-3mm}\\\vspace{-1mm}%
    \begin{list2}
        \item Thesis Advisor:  Dr. William F. Punch
        \item GPA: 4.0/4.0
        \item ABD, Expected Graduation: June 2015
    \end{list2}\vspace{-1.5mm}
    \textbf{Missouri S\&T}, Rolla, Missouri \vspace{2mm}\\\vspace{1mm}%
    \textsl{Master of Science in Computer Science} \hfill \textbf{ August 2010 -- May 2012}\vspace{-3mm}\\\vspace{-1mm}%
    \begin{list2}
        \item Thesis Advisor:  Dr. Daniel R. Tauritz
        \item GPA: 4.0/4.0
        \item Summa Cum Laude
    \end{list2}\vspace{-1.5mm}
    \textsl{Bachelor of Science in Computer Science} \hfill \textbf{ August 2006 -- July 2010}\vspace{-3mm}\\\vspace{-1mm}%
    \begin{list2}
        \item Minor in Mathematics
        \item GPA: 3.84/4.0 in major, 3.81/4.0 overall
        \item Summa Cum Laude
    \end{list2}\vspace{-1.5mm}

    %__________________________________________________________________________________________________________________
    % Research
    \section{\mysidestyle Peer Reviewed\\Journal Articles}

    \begin{enumerate}
    \item
    B. W. Goldman and W. F. Punch,
    ``Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid,''
    \textbf{Invited Paper}, \textsl{Evolutionary Computation Journal}, In Review

    \item
    B. W. Goldman and W. F. Punch,
    ``Analysis of Cartesian Genetic Programming's Evolutionary Mechanisms,''
    \textsl{IEEE Transactions on Evolutionary Computation}, In Press

    \item
    K. D. S. Lehmann,  B. W. Goldman, I. Dworkin, D. M. Bryson, A. P. Wagner,
    ``From Cues to Signals: Evolution of Interspecific Communication via Aposematism and Mimicry in a Predator-Prey System,''
    \textsl{PloS one}, vol. 9, no. 3, e91783, 2014.

    \item
    D. R. White, J. McDermott, M. Castelli, L. Manzoni, B. W. Goldman,
    G. Kronberger, W. Jaskowski, U.-M. O’Reilly, and S. Luke,
    ``Better GP Benchmarks: Community Survey Results and Proposals,''
    \textsl{Genetic Programming and Evolvable Machines}, vol. 14, no. 1, pp. 3--29, 2013.    
    \end{enumerate}

    \section{\mysidestyle Peer Reviewed\\Conference Papers}
    \begin{enumerate}

    \item
    B. W. Goldman and W. F. Punch,
    ``Parameter-less Population Pyramid,''
    \textsl{GECCO}, pp. 785--792, 2014
    \textbf{Won Best Paper in Track}

    \item
    B. W. Goldman and W. F. Punch,
    ``Length Bias and Search Limitations in Cartesian Genetic Programming,''
    \textsl{GECCO}, pp. 933--940, 2013

    \item
    N. R. Kamrath, B. W. Goldman, and D. R. Tauritz,
    ``Using Supportive Coevolution to Evolve Self-Configuring Crossover,''
    \textsl{GECCO}, pp. 1489--1496, 2013.

    \item
    B. W. Goldman and W. F. Punch,
    ``Reducing Wasted Evaluations in  Cartesian Genetic Programming,''
    \textsl{EuroGP}, pp. 61--72, 2013

    \item
    B. W. Goldman and D. R. Tauritz,
    ``Linkage tree genetic algorithms: variants and analysis,''
    \textsl{GECCO}, pp. 625--632, 2012

    \item
    B. W. Goldman and D. R. Tauritz,
    ``Supportive Coevolution,''
    \textsl{GECCO}, pp. 59--66, 2012.

    \item
    B. W. Goldman and D. R. Tauritz,
    ``Self-Configuring Crossover,''
    \textsl{GECO}, pp. 575--582, 2011.
    \end{enumerate}
    \section{\mysidestyle Peer Reviewed\\Poster Papers}
    \begin{enumerate}
    \item
    B. W. Goldman and D. R. Tauritz,
    ``Meta-Evolved Empirical Evidence of the Effectiveness of Dynamic Parameters'',
    \textsl{GECCO}, pp. 155--156, 2011.
    \end{enumerate}
    \pagebreak

    %__________________________________________________________________________________________________________________
    % Teaching Experience
    \section{\mysidestyle Teaching\\Experience}
    \textbf{Michigan State University}, East Lansing, Michigan \vspace{2mm}\\\vspace{1mm}%
    \textsl{CSE232: Introduction to Programming II} \hfill \textbf{May 2014 -- July 2014}\\
    Lead instructor for course. In charge of lecturing, creating course content, and assessments
    for 37 students. Managed two graduate assistants who taught lab sections and graded assignments.
    Received 3.81/4 student evaluation.
    Topics include the C++ programming language from basics up to dynamic memory, templating, and user created data structures.

    \textsl{CSE232: Introduction to Programming II} \hfill \textbf{August 2013 -- May 2014}\\
    Primary instructor for lab section,
    topics include the C++ programming language from basics up to dynamic memory, templating, and user created data structures.

    \textbf{Missouri S\&T}, Rolla, Missouri \vspace{2mm}\\\vspace{1mm}%
    \textsl{CS328: Object Oriented Numerical Methods} \hfill \textbf{January 2012 -- May 2012}\\
    Grader for intensive C++ course,
    topics include advanced programming techniques and how to efficiently solve large systems of equations.

    \textsl{CS54: Introduction to C++ Lab} \hfill \textbf{January 2011 -- December 2011}\\
    Primary instructor for introductory computer science course,
    topics include programming basics up to inheritance, style, editors, and how to interact with Unix.

    \textsl{CS387: Parallel Computing} \hfill \textbf{January 2011 -- May 2011}\\
    Teaching assistant for parallel computation course,
    topics include MPI, cluster computing, GPUs, Monte Carlo experiments, and high performance computing.

    \textsl{CS348: Evolutionary Computation} \hfill \textbf{August 2010 -- December 2010}\\
    Teaching assistant for evolutionary computation course,
    topics include genetic algorithms, evolutionary strategies, genetic programming, and learning classifier systems.

    %__________________________________________________________________________________________________________________
    % Professional Experience
    \section{\mysidestyle Recent\\Professional\\Experience}

    \textbf{Los Alamos National Laboratories}, Los Alamos, New Mexico \vspace{2mm}\\\vspace{1mm}%
    \textsl{Advanced Computing Solutions Program} \hfill \textbf{May 2012 -- August 2012}\\
    Investigated intrusion detection techniques using automated network behavior analysis.

    \textbf{Sandia National Laboratories}, Albuquerque, New Mexico \vspace{2mm}\\\vspace{1mm}%
    \textsl{Center for Cyber Defenders Graduate Student Intern} \hfill \textbf{May 2011 -- August 2011}\\
    Developed covert command and control communication methods, created security audit tools for a
    surveillance camera system, and designed an educational game to teach secure system design.

    \textbf{Dynetics}, Huntsville, Alabama \vspace{2mm}\\\vspace{1mm}%
    \textsl{Computer Analyst 3} \hfill \textbf{May 2010 -- August 2010}\\
    Worked with department of defense applications to improve unmanned aerial vehicle data analysis.

    \begin{comment}
    \textbf{AT\&T}, Saint Louis, Missouri \vspace{2mm}\\\vspace{1mm}%
    \textsl{Summer Technical Intern} \hfill \textbf{May 2008 -- August 2008}\\
    Developed automated service detection to ensure all pertinent applications installed on a server
    were restarted at server reboot.  Modernized legacy software performing phone number database queries.
    \end{comment}

    %__________________________________________________________________________________________________________________
    % Honors and Awards
    \section{\mysidestyle Honors and\\Activities}
    Best Paper - Genentic Algorithms Track for GECCO 2014 ``Parameter-less Population Pyramid''\\%
    Peer Reviewer, IEEE Transactions on Evolutionary Computation Journal 2014--present\\%
    Founder and organizer, Computational Evolution Discussion Group October 2013--present\\%
    Program Committee Member, GECCO 2012--present Genetic Algorithms Track \vspace{1mm}\\%
    Voted 2012 Leader of the Year, Missouri S\&T Computer Science Department \vspace{1mm}\\%
    %Chair of ACM SIG for AI competition game development (SIG-Game), 2011--2012 \vspace{1mm}\\%
    Google AI challenge, ranked 25$^{th}$ out of 4619 world wide, 6$^{th}$ in USA, 2010 \vspace{1mm}\\%
    Missouri S\&T Human versus Computer chess tournament, two first place AIs, 2009 and 2011 \vspace{1mm}\\%
\begin{comment}    
    ACM SIG-Game Competition testing and balancing using evolving AI, 2009--2012 \vspace{1mm}\\%
    ACM SIG-Game first place AI, 2008          \vspace{1mm}\\%
    University of Illinois at Urbana-Champaign MechMania AI competition second place AI, 2007  \vspace{1mm}\\%
    Missouri S\&T Alumni Scholarship 2006--2010             %\vspace{1mm}\\%

    %__________________________________________________________________________________________________________________
    % Computer Skills
    \section{\mysidestyle Programming} 
    Python, C, C++, C\#, Matlab, Linux shell scripting, \LaTeX, Java, CUDA.

%______________________________________________________________________________________________________________________
\section{\mysidestyle References}

\begin{tabular}{@{}p{6cm}p{6cm}}
\textbf{Dr. Daniel R. Tauritz} (Advisor)       &  \textbf{Professor Fikret Ercal}                   \\
Associate Professor                &  Professor                       \\
Department of Computer Science     & Department of Computer Science \\
Missouri S\&T                        &  Missouri S\&T                      \\
Rolla, Missouri                      &  Rolla, Missouri        \\
phone: \textsl{(573) 341-7218}    &  phone: \textsl{(573) 341-4857}     \\
e-mail: \textsl{tauritzd@mst.edu}   &  e-mail: \textsl{ercal@mst.edu}    \\
\end{tabular}

\begin{tabular}{@{}p{6cm}p{6cm}}
\textbf{Mr. Dan Thomsen}                \\%&  \textbf{Mr Sean Loye}                    \\
Senior Scientist                        \\      %&  Systems Engineer                         \\
Sandia National Laboratories            \\      %&  Hewlett Packard                          \\
Albuquerque, New Mexico                 \\ %&  Milton, Queensland, Australia            \\
phone: \textsl{(612) 789-0559}    \\%&  phone: \textsl{available on request}     \\
e-mail: \textsl{available on request}   \\%&  e-mail: \textsl{available on request}    \\
\end{tabular}

\end{comment}
%______________________________________________________________________________________________________________________
\end{resume}
\end{document}

%______________________________________________________________________________________________________________________
% EOF