### 2015-03-03 Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid/reviews2.tex
Adding the second response to the editor.

\documentclass[]{article}
\usepackage{hyperref}

\begin{document}

\begin{quote}
So, I would leave it with the authors. If they are willing to include results on hBOA and/or LTGA with local search, that’s fantastic.
If they conclude that they would rather not spend the extra effort, then I would ask them to at least comment on the issue in the paper.
\end{quote}

\begin{quote}
You could formulate it very cautiously, clearly saying that these are just preliminary observations and that on some problems local search does help (MAXSAT).
\end{quote}

In Section 5 Comparison Algorithm Parameter Tuning, paragraph at the start of page 13: We now explain how
we performed preliminary tests of using local search with hBOA and LTGA, but found it to generally not helpful. We
highlight the three cases in which local search improved results.

The only other modification to the paper was in Figure 4, top of page 9: Line 5 and 11 now indicate that a specific
level's model is being rebuilt, clarifying that there is a different model for each level.
\end{document}