### 2015-04-15 Dissertation/chapter-understanding.tex
New chapter stuff

\chapter{Using Gray-Box Methods to understand Black-Box Effectiveness}

\section{Big Valley}

\section{Pyramid Levels}