### 2015-06-24 Dissertation/chapter-conclusions.tex

\chapter{Conclusions and Future Work}
% P3 Journal
The Parameter-less Population Pyramid (P3) is a recently introduced method for performing black-box
optimization. P3's primary innovation is the replacement of the generational model with a pyramid of populations.
This pyramid is constructed iteratively, with both the number of levels and the number of solutions stored
at each level growing as search progresses. P3 uses a model based crossover method
that learns a linkage tree from gene epistasis. Combined with a simple hill climber, P3's design contains
many synergistic features.

Across a large number of problems and problem sizes P3 required fewer evaluations to reach the global optimum
than optimally tuned state-of-the-art competitors. On single-instance problems P3's improvement was by a
constant factor, while for the three randomly generated problem classes P3's improvement increased with problem size. This
quality extends to intermediate points during evolution, with P3 generally reaching at least as high
a fitness as the competitive techniques when using the same number of evaluations. While P3 does require
modeling overhead, the expense of this overhead is approximately linear with respect to genome size. There
is some evidence that even when compared on wall clock time, P3 performs on par with
the best comparison techniques. All of these achievements are made without any problem-specific parameter
tuning, making P3 easier to apply to new domains than its two closest competitors in quality.

P3's quality is due to a number of desirable traits. First, mixing local search with model based
crossover lets search focus on properly mixing high-quality solutions. Second, by adding diversity only
as necessary P3 tends to use the minimal amount of random initialization, unlike other techniques that must
overcompensate with larger population sizes on single-instance problems and consider the worst instance
when solving problem classes. Third, by heavily exploiting existing diversity before adding more P3 is able
to reach high-quality intermediate fitnesses quickly without prematurely converging. Fourth, the
nature of the pyramid's shape allows search to preserve a desirable proportion of diversity at
each fitness level, similar to a generational model using a decreasing population size.

When leveraging features of the new domain, Gray-Box P3 can obtain substantial improvement
over Black-Box P3. This comes from the integration of the Hamming-Ball Hill Climber (HBHC)
and a novel method for deriving linkage clusters from the known problem epistasis,
both of which result in a factor $N$ speedup over their black-box counterparts.
Furthermore, Gray-Box P3
can perform partial reevaluation and efficient hill climbing during the mixing phase
without significant overhead. In specializing, Gray-Box P3 is able to find
global optima orders of magnitude faster than Black-Box P3.

There are a number of meaningful avenues for future P3 experimentation. Perhaps the most
pressing for practitioner acceptance is to apply P3 to real-world problems and compare
its results with other black-box or even problem-specific heuristics. While parameter-less,
P3 is currently limited to discrete, fixed-length genomes evaluated using single-objective
fitness. These limitations can be relaxed with future work to make P3 more widely applicable.
While asymptotically linear in problem size, Black-Box P3's modeling techniques and local search methods are likely going
to be prohibitively expensive for genome sizes in the hundreds of thousands or millions of genes,
and the inability of the model to capture overlapping linkage may be hindering
search efficiency. Overcoming these limitations by using a new modeling technique may allow
the pyramid model even greater flexibility. Similarly, while P3 is able to overcome low-order
deception via linkage learning, the iterative improvement method by which crossovers
are made may mislead search on landscapes with higher-order deception. This appears
to be especially important on NP-Hard problems, where high-quality solutions may be
very distant from the global optimum in representation space.

While empirically effective, we believe it is possible to derive theoretical foundations
for Gray-Box P3. Specifically, this may be possible by considering the nature of crossover
in terms similar to those used for finding $r$-bit local optima with HBHC. Consider
that for any two parents the non-linear effects of crossover are constrained by
where they contain different bit values. As a result, subsets of their differences
can become linearly separable, allowing for each to be considered independently of all others.
Furthermore, these conclusions about effective crossover in gray-box can likely be translated
into a black-box setting to help explain the effectiveness of Black-Box P3.

However, even without
these improvements our results show P3 is highly efficient at finding
global optima on both black-box and gray-box problems without any problem-specific tuning.