### 2015-10-27 CV/Brian Goldman CV.tex
Updated CV from writing group feedback

\documentclass[a4paper,11pt]{article}
\usepackage{cv}
\usepackage{verbatim}
\name{Brian W. Goldman}
\info{Address: & Department of Computer Science \& Engineering, Michigan State University, East Lansing, Michigan.\\
      Phone: & 314-313-1281\\
      Email: & brianwgoldman@acm.org\\
      Website: & \url{http://brianwgoldman.github.io/}}

\bibliography{../main}

\addtocategory{journals}{goldman:2015:cgpanalysis,goldman:2015:fastp3,white:2013:bgpb,lehmann:2013:mimics}
%At MSU
\addtocategory{papers}{goldman:2014:p3,goldman:2015:GBO}
%CGP
\addtocategory{papers}{goldman:2013:ordering,goldman:2013:cgpwaste}
%With Tauritz
\addtocategory{papers}{goldman:2012:ltga,goldman:2011:scx,kamrath:2013:supportivescx,goldman:2012:supportive}

\addtocategory{posters}{goldman:2011:dynamic-parameters}

\addtocategory{curpapers}{goldman:2016:hyperplane,goldman:2016:p3hiff,whitley:2016:mkl}
%\ref{sumpapers} A list of these appears on pages \pageref{papersstart}--\pageref{papersend}.

\begin{document}
\maketitle

\section{Education}
\begin{itemize}
\item 2015 Doctor of Philosophy in Computer Science \& Engineering, Michigan State University
\begin{itemize}
  \item Dissertation Advisor:  Dr. William F. Punch%, GPA: 4.0/4.0
  \item Dissertation: Out of the Box Optimization using the Parameter-less Population Pyramid
\end{itemize}
\item 2012 Master of Science in Computer Science, Missouri University of Science \& Technology
\begin{itemize}
  \item Thesis Advisor:  Dr. Daniel R. Tauritz%, GPA: 4.0/4.0
  \item Thesis: Robust Evolutionary Algorithms
\end{itemize}
\item 2010 Bachelor of Science in Computer Science, Missouri University of Science \& Technology
%\begin{itemize}
%  \item Minor in Mathematics, GPA: 3.84/4.0 in major, 3.81/4.0 overall
%\end{itemize}
\end{itemize}
\section{Academic Positions}
\begin{itemize}
\item Research Assistant Professor, Colorado State University. Starting Spring 2016.
\item Postdoc, Hintze Lab, Michigan State University. Fall 2015.
\item Advanced Computing Solutions Program, Los Alamos National Laboratories. Summer 2012.
\item Center for Cyber Defenders, Sandia National Laboratories. Summer 2011.
\end{itemize}

\begin{publications}
%\printbib{books}
\printbib{journals}
\printbib{papers}
\printbib{posters}
\printbib{curpapers}
%\printbib{chapters}
%\printbib{techreports}
%\printbib{bookreviews}
%\printbib{editorials}
\end{publications}

\section{Grants}
\begin{itemize}
\item BEACON NSF STC 2015 funded proposal \$105,148
\item BEACON NSF STC 2013 funded proposal \$62,284.
\end{itemize}

\section{Awards}
\begin{itemize}
\item     Runner-up, Best Paper - Genetic Algorithms Track for GECCO 2015 ``Gray Box Optimization using the Parameter-less Population Pyramid''
\item     Best Paper - Genetic Algorithms Track for GECCO 2014 ``Parameter-less Population Pyramid''
\item     Voted 2012 Leader of the Year, Missouri S\&T Computer Science Department
\item     Google AI challenge, ranked 25$^{th}$ out of 4619 world wide, 6$^{th}$ in USA, 2010
\item     Missouri S\&T Human versus Computer chess tournament, two first place AIs, 2009 and 2011
\end{itemize}

\section{Invited talks}
\begin{itemize}
\item Departmental Seminar, University of Memphis, April 2015
\item BEACON NSF STC Congress, Michigan State University, August 2014 and August 2015
\item Evolutionary Computation, Michigan State University, September 2014 and October 2015
\item Heuristic and Evolutionary Algorithms Laboratory, University of Applied Sciences Upper Austria, December 2014
\item BEACON Day, North Carolina A\&T, October 2013
\end{itemize}

\section{Activities}
\begin{itemize}
\item     Co-chair, 2016 Genetic Programing Theory and Practice Conference
\item     Co-chair, GECCO 2015--present Combinatorial Black Box Optimization Competition
\item     Peer Reviewer, IEEE Transactions on Evolutionary Computation Journal 2014--present
\item     Peer Reviewer, Genetic Programming and Evolvable Machines 2014--present
\item     Program Committee Member, GECCO 2012--present Genetic Algorithms Track
\item     Founder and organizer, Computational Evolution Discussion Group October 2013--2015
\item     Chair of ACM SIG for AI competition game development (SIG-Game), 2011--2012
\end{itemize}

\section{Teaching Experience}
\begin{itemize}
\item Introduction to Programming II, Summer 2014.
      Lead instructor for C++11 course. In charge of lecturing, creating course content, and assessments
      for 37 students. Managed two graduate assistants who taught lab sections and graded assignments.
      Received 3.81/4 student evaluation.
\item Introduction to Programming II, Fall 2013 -- Spring 2014. Primary instructor for lab section.
\item Object Oriented Numerical Methods, Spring 2012. Teaching assistant for upper level intensive C++ course.
\item Introduction to C++ Lab, Spring 2011 -- Fall 2011. Primary instructor for lab section.
\item Parallel Computing, Spring 2011. Teaching assistant for upper level course.
\item Evolutionary Computation, Fall 2010. Teaching assistant for upper level course.
\end{itemize}

\end{document}